The Spine of American Law: Digital History Methods and U.S. Legal Practice

### 2016-07-28 article/Funk-Mullen.Spine-of-Legal-Practice.Rmd
Paragraph on digital textual analysis

Yet by turning to the digital analysis of texts, we have resolved this difficulty and tracked how states borrowed their codes of civil practice from one another. Within the corpus of legislation, algorithmic analysis of texts can reverse engineer and visualize what the archive revealed in figure 1: which texts were borrowed, which modified, and how extensively. Our method works especially well for legal texts, for reasons we will explain. However, scholars can extend this method to other curated collections of texts, from church hymns to medical treatises.

@@

@@

[^31]: We have preferred to use the term "digital history" when referring to our own work, in part because the term "digital humanities" has largely come to refer to the work of digital literary and digital media scholars, but primarily because we wish to see digital scholars make disciplinary, rather than interdisciplinary contributions. We are working primarily in the field formerly known as humanities computing, but there are other forms of digital history such as digital public history or spatial history. On the role of disciplines and the importance of field specific argumentation, see, Stephen Robertson, "The Differences between Digital Humanities and Digital History," 289--307, and Cameron Blevins, "Digital History's Perpetual Future Tense," 308--324, both in *Debates in the Digital Humanities 2016*, ed. Matthew K. Gold and Lauren F. Klein (Minneapolis: University of Minnesota Press, 2016); William G. Thomas III, "The Promise of the Digital Humanities and the Contested Nature of Digital Scholarship," in *A New Companion to the Digital Humanities*, edited by Susan Schreibman, Ray Siemens, and John Unsworth (Malden, MA: Wiley Blackwell, 2016), 524--37. Nevertheless, digital text analysis in the humanities has mostly been published by literary scholars, including Stephen Ramsay, *Reading Machines: Toward an Algorithmic Criticism* (Urbana: University of Illinois Press, 2011); Franco Moretti, *Distant Reading* (New York: Verso, 2013); Matthew L. Jockers, *Macroanalysis: Digital Methods and Literary History* (Urbana: University of Illinois Press, 2013); Ted Underwood, *Why Literary Periods Mattered: Historical Contrast and the Prestige of English Studies* (Stanford: Stanford University Press, 2013); Ryan Cordell and David A. Smith, *Viral Texts: Mapping Networks of Reprinting in 19th-Century Newspapers and Magazines*, NULab for Texts Maps and Networks, Northeastern University (2012--): <http://viraltexts.org>. Related but less well regarded by humanities scholars is work in "culturomics": see Jean-Baptiste Michel et al., "Quantitative Analysis of Culture Using Millions of Digitized Books", *Science* 331, no. 6014 (2011), 176--182; doi:10.1126/science.1199644. On the uselessness of the term "big data," see Ted Underwood, "Against (Talking About) 'Big Data,'" *The Stone and the Shell*, blog post, 10 May 2013: <https://tedunderwood.com/2013/05/10/why-it-matters-that-we-dont-know-what-we-mean-by-big-data/>. For an overview of digital history projects involving text analysi, see the Roberston essay cited above. For an example of curating a corpus aimed at research questions, see Ted Underwood, Boris Capitanu, Peter Organisciak, Sayan Bhattacharyya, Loretta Auvil, Colleen Fallaw, J. Stephen Downie, "Word Frequencies in English-Language Literature, 1700--1922," dataset, v0.2 (HathiTrust Research Center, 2015) doi:10.13012/J8JW8BSJ.

>>>>

Yet by turning to the digital analysis of texts, we have resolved this difficulty and tracked how states borrowed their codes of civil practice from one another. Within the corpus of legislation, algorithmic analysis of texts can reverse engineer and visualize what the archive revealed in figure 1: which texts were borrowed, which modified, and how extensively. Our method works especially well for legal texts, for reasons we will explain.

The computational analysis of texts is a method which historians can use across the discipline to study many topics. We have used text analysis---specifically, methods to detect text reuse---to show how law migrates through borrowings, just as scholars on the *Viral Texts* team have demonstrated how newspaper articles were reprinted in the nineteenth-century United States.[^100] While we have developed a method that discovers borrowings of exact words and phrasings, other forms of digital text analysis can track the diffusion of concepts. Making due allowance for the specific historical questions and sources at hand, computational text analysis is broadly applicable to historical problem that involves the spread of words or ideas. The specific method that we outline could be extended beyond codes of civil procedure to legal statutes or treatises. Yet it could also be used, for instance, to track the spread of hymns in collections of hymnbooks in religious history, or the reuse of sections of medical textbooks in the history of science. Since historians by and large work with textual sources and increasingly with digitized texts, computational text analysis should become a part of the historian's toolbox. Lara Putnam has recently described the importance of "digitization and OCR, which make words above all available" for historians to search and read, while observing that "computational tools can discipline our term-searching if we ask them to."[^101] Historians have been slow to take up text analysis even as the aid to more traditional reading that Putnam recommends. Yet as we demonstrate in this article, such methods can reveal patterns inaccessible to the traditional historian.[^102]

@@

[^100]: Ryan Cordell and David A. Smith, *Viral Texts: Mapping Networks of Reprinting in 19th-Century Newspapers and Magazines*, NULab for Texts Maps and Networks, Northeastern University (2012--): <http://viraltexts.org>.

[^101]: Lara Putnam, "The Transnational and the Text-Searchable: Digitized Sources and the Shadows They Cast," *American Historical Review* 121, no. 2 (2016): 399--400.

[^102]: Only a few historians have used computational text analysis, including Cameron Blevins, “Space, Nation, and the Triumph of Region: A View of the World from Houston,” *Journal of American History* 101, no. 1 (June 1, 2014): 122–47, doi:10.1093/jahist/jau184; Benjamin M. Schmidt, *Sapping Attention* (blog): <http://sappingattention.blogspot.com/>; Sharon Block, "Doing More with Digitization: An Introduction to Topic Modeling of Early American Sources," *Common-Place* 6, no. 2 (January 2006): <http://www.common-place.org/vol-06/no-02/tales/>; Dan Cohen, Frederick Gibbs, Tim Hitchcock, Geoffrey Rockwell, et al., "Data Mining with Criminal Intent," white paper, 31 August 2011, <http://criminalintent.org>; Robert K. Nelson, "Mining the Dispatch, website, Digital Scholarship Lab, University of Richmond, <http://dsl.richmond.edu/dispatch/>; Dan Cohen, "Searching for the Victorians," 4 October 2010: <http://www.dancohen.org/2010/10/04/searching-for-the-victorians/>; Micki Kaufmann, *"Everything on Paper Will Be Used Against Me:" Quantifying Kissinger*, digital project (2012--16): <http://blog.quantifyingkissinger.com/>; E. Thomas Ewing, Samah Gad, Bernice L. Hausman, Kathleen Kerr, Bruce Pencek, and Naren Ramakrishnan, "An Epidemiology of Information: Datamining the 1918 Flu Pandemic," project research report (2 April 2014): <http://vtechworks.lib.vt.edu/bitstream/handle/10919/46991/An%20Epidemiology%20of%20Information%20Project%20Research%20Report_Final.pdf?sequence=1>; Michelle Moravec, "'Under this name she is fitly described': A Digital History of Gender in the History of Woman Suffrage" (March 2015): <http://womhist.alexanderstreet.com/moravec-full.html>.

@@

[^31]: We have preferred to use the term "digital history" when referring to our own work, in part because the term "digital humanities" has largely come to refer to the work of digital literary and digital media scholars, but primarily because we wish to see digital scholars make disciplinary, rather than interdisciplinary contributions. We are working primarily in the field formerly known as humanities computing, but there are other forms of digital history such as digital public history or spatial history. On the role of disciplines and the importance of field specific argumentation, see, Stephen Robertson, "The Differences between Digital Humanities and Digital History," 289--307, and Cameron Blevins, "Digital History's Perpetual Future Tense," 308--324, both in *Debates in the Digital Humanities 2016*, ed. Matthew K. Gold and Lauren F. Klein (Minneapolis: University of Minnesota Press, 2016); William G. Thomas III, "The Promise of the Digital Humanities and the Contested Nature of Digital Scholarship," in *A New Companion to the Digital Humanities*, edited by Susan Schreibman, Ray Siemens, and John Unsworth (Malden, MA: Wiley Blackwell, 2016), 524--37. Nevertheless, digital text analysis in the humanities has mostly been published by literary scholars, including Stephen Ramsay, *Reading Machines: Toward an Algorithmic Criticism* (Urbana: University of Illinois Press, 2011); Franco Moretti, *Distant Reading* (New York: Verso, 2013); Matthew L. Jockers, *Macroanalysis: Digital Methods and Literary History* (Urbana: University of Illinois Press, 2013); Ted Underwood, *Why Literary Periods Mattered: Historical Contrast and the Prestige of English Studies* (Stanford: Stanford University Press, 2013). Related but less well regarded by humanities scholars is work in "culturomics": see Jean-Baptiste Michel et al., "Quantitative Analysis of Culture Using Millions of Digitized Books", *Science* 331, no. 6014 (2011), 176--182; doi:10.1126/science.1199644. On the uselessness of the term "big data," see Ted Underwood, "Against (Talking About) 'Big Data,'" *The Stone and the Shell*, blog post, 10 May 2013: <https://tedunderwood.com/2013/05/10/why-it-matters-that-we-dont-know-what-we-mean-by-big-data/>. For an overview of digital history projects involving text analysi, see the Roberston essay cited above. For an example of curating a corpus aimed at research questions, see Ted Underwood, Boris Capitanu, Peter Organisciak, Sayan Bhattacharyya, Loretta Auvil, Colleen Fallaw, J. Stephen Downie, "Word Frequencies in English-Language Literature, 1700--1922," dataset, v0.2 (HathiTrust Research Center, 2015) doi:10.13012/J8JW8BSJ.
